---
title: "Intake Simulation Summary"
description: "Summary of a test grid simulation for jail intake"
author:
  - name: "Krzysztof Sakrejda"
  - url: "https://github.com/epibayes/jail-model"
  - affiliation: "Epibayes"
  - affiliation_url: "https://epibayes.io"
date: "`r Sys.Date()`"
output: distill::distill_article
params: 
  simulation_name: "default-test"
  simulation_id: "00001"
  replicate_id: "001"
---

```{r setup, include=FALSE}
library(magrittr)
knitr::opts_chunk$set(echo = FALSE)
simulation_name = params$simulation_name
simulation_id = params$simulation_id
replicate_id = params$replicate_id
```

```{r model-parameters, include=FALSE}
parameters = workflow::build_dir(simulation_name, "parameter-grid.rds") %>% 
  readRDS() %>% dplyr::filter(
    simulation_name == simulation_name,
    simulation_id == simulation_id,
    replicate_id == replicate_id)
rmarkdown::paged_table(parameters)
```

The intake simulation follows the intake diagram, with the exception that the
mixed RTU cohort is not currently implemented.  

```{r intake-diagram, layout="l-body-outset"}
intake_fig_path = workflow::project_dir("static", "diagrams", "covid-jail-intake-diagram.svg")
```

Incoming detainees are administered
a rapid COVID-19 test, followed by an X-ray screening for tuberculosis which can
uncover both symptomatic and asymptomatic COVID-19 cases.  Individuals testing
positive for COVID-19 or with abnormal chest X-rays images are effectively
isolated (COVID-19+ together, and those with chest X-ray abnormalities are
"Persons Under Investigation" (PUI).  Those with normal results are placed in
daily cohorts for 14 days, retested, and then released into the general
population.  Individuals testing negative with additional medical needs enter
the Residential Treatment Unit (RTU) intake cohort for 14 days and are not
cohorted into daily groups.

![Diagram of the jail intake process. The three intake cohorts are COVID-19+
(or with abnormal chest X-rays, although those are isolated separately), daily
COVID-19 negative cohorts with daily symptomatic screening, and the mixed
Residential Treatment Unit (RTU) intake cohort.](`r intake_fig_path`){width=100%}

After 14 days, we assume all individuals are screened for COVID-19 via a PCR
test and upon a negative test they are admitted to the general population.
Those who are screened and test positive at any point in time join the COVID-19
positive cohort until they are medically cleared to join the general
population.  Those in the mixed RTU cohort enter the general RTU population
after 14 days if negative for COVID-19 on a PCR test.

We summarize the number of cohorts with undetected cases at the 14 day mark, 
labelled as "breakthrough" cases.  Their PMF and CMF are summarized below as
the percentage of cohorts in each category.

```{r burden-pmf, layout='l-body-outset'}
burden_pmf = workflow::artifact_dir(simulation_name, simulation_id, 
  replicate_id, "breakthrough-cases-pmf.svg") %>% fs::path_expand()
```

![Percentage of cohorts with M undetected cases.  Most cohorts are COVID-19
negative but a significant proportion have 1+ breakthrough cases.](`r burden_pmf`){width=100%}


```{r burden-cmf, layout='l-body-outset'}
burden_cmf = workflow::artifact_dir(simulation_name, simulation_id, 
  replicate_id, "breakthrough-cases-cmf.svg") %>% fs::path_expand()
```

![Percentage of cohorts with at least M undetected cases.  Most cohorts include
no COVID-19 positive individuals, so effects on general population epidemics
are driven by a small number of cases that break through the cohorting 
procedure.](`r burden_cmf`){width=100%}

```{r burden-over-time, layout='l-body-outset'}
burden_time = workflow::artifact_dir(simulation_name, simulation_id, 
  replicate_id, "breakthrough-cases-case-percent.svg") %>% fs::path_expand()
```
 
![The frequency of cohorts with, 1, 2, 3, or 4+ cases.  Over the first 14 days
the frequency of cohorts with nonzero cases drops. At day 14 after PCR 
testing the number of cohorts with 1 breakthrough case rises as cases are 
removed from higher-burden cohorts.  
](`r burden_time`){width=100%}


```{r overall_trace, layout='l-body-outset'}
overall_trace = workflow::artifact_dir(simulation_name, simulation_id, 
  replicate_id, "summary-trace-plot.svg") %>% fs::path_expand()
```
 
![Incidence of new cases is distributed over much of the cohorting
period as exposed but non-infectious individuals can not be detected in initial 
screening.  Secondary cases can develop over much of the cohorting period and 
the average number of undetected cases is relatively consistent during this
period.  At day 14 more cases are detected and isolated but which reduced
but does not eliminate breakthrough cases.](`r overall_trace`){width=100%}
