---
title: "Cook County Correctional Facility Data Summary"
description: "Summary of administrative and movement data."
author:
  - name: "Krzysztof Sakrejda"
  - url: "https://github.com/epibayes/jail-model"
  - affiliation: "Epibayes"
  - affiliation_url: "https://epibayes.io"
date: "`r Sys.Date()`"
output: distill::distill_article
params: 
  standard_capacity_data: "tibble::tibble()"
  standard_occupancy_data: "tibble::tibble()"
  duration_of_detention_quantiles: "tibble::tibble()"
  standard_count_data: "tibble::tibble()"
  applied_test_sensitivity: "0"
  building_count: "tibble::tibble()"
  building_percentage: "tibble::tibble()"
  unit_transition_plot: "ggplot()"
  unit_mmo: "tibble::tibble()"
---

```{r setup, include=FALSE}
library(magrittr)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE)
```

Our simulations are based on summaries that represent the processes in
place at the Cook County Department of Corrections site in Chicago Illinois.  
We combined administrative data with summaries of past studies and 
individual correspondence to characterize the intake of detainees, 
movement within the facility at the unit level, timing of movement, 
timing of release, and facility structure to produce this overview.

# Facility 

```{r capacity, include = FALSE}
n_units = params$standard_capacity_data %>% nrow()
building_capacity = params$standard_capacity_data %>% 
  dplyr::group_by(building, type) %>% 
  dplyr::summarise(
    `# of units` = dplyr::n(),
    `housing capacity` = capacity %>% as.numeric() %>% sum)
```

We focused on the main living units on the CCDC campus, including seven celled
units, three dorms, a residential treatment unit (RTU), and the single "Boot
Camp" unit.

```{r building-capacity-chart, include = TRUE}
#rmarkdown::paged_table(building_capacity, 
#  options = list(rows.print = 20, paged.print = FALSE))
knitr::kable(building_capacity, caption = 'Building units and capacity')
```

# Intake Process

```{r intake-diagram, layout="l-body-outset"}
intake_fig_path = workflow::project_file("static", "diagrams", "covid-jail-intake-diagram.svg")
```

Incoming detainees are administered
a rapid COVID-19 test, followed by an X-ray screening for tuberculosis which can
uncover both symptomatic and asymptomatic COVID-19 cases.  Individuals testing
positive for COVID-19 or with abnormal chest X-rays images are effectively
isolated.  Individuals who are COVID-19+ a single COVID-19+ cohort, and those with 
chest X-ray abnormalities are classified as "Persons Under Investigation" (PUI)
and separated individually until that status is resolved.  Those with normal 
results are transferred to daily cohorts for 14 days, retested, and then 
released into the general population once they are cleared by [WHO?].  Individuals 
testing negative with additional medical needs enter the Residential Treatment 
Unit (RTU) intake cohort for 14 days and are not cohorted into daily groups.

![Figure 1: Diagram of the jail intake process. The three intake cohorts are COVID-19+
(or with abnormal chest X-rays, although those are isolated separately), daily
COVID-19 negative cohorts with daily symptomatic screening, and the mixed
Residential Treatment Unit (RTU) intake cohort.](`r intake_fig_path`){width=100%}

```{r blank, fig.caption = "", include = FALSE}
```

Typically between days 10 and 14 days all individuals are screened for COVID-19 
via a PCR test and upon a negative test they are admitted to the general population.
Those who are screened and test positive at any point in time join the COVID-19
positive cohort until they are medically cleared to join the general
population.  Those in the mixed RTU cohort enter the general RTU population
after 14 days if negative for COVID-19 on a PCR test.

# Intake of detainees and duration of detention


```{r intake-data, include = FALSE}
iqr = quantile(params$standard_count_data$intake, probs = c(0.25, 0.75), na.rm = TRUE)
l1_pacf = params$standard_count_data$intake %>% 
  purrr::discard(is.na) %>% 
  pacf(lag.max = 10) %>% 
  `[[`('acf') %>% `[`(1,1,1) %>%
  round(2)
```

Daily detainee intake count is relatively consistent, has an 
IQR of [`r paste0(iqr[1], ", ", iqr[2])`] and shows positive 
autocorrelation (partial autocorrelation of `r l1_pacf` at lag 1).

```{r intake-time-series, layout="l-body-outset", include = TRUE, fig.cap = "Daily detainee intake count for the period covered by the data and the corresponding histogram of daily intake counts."} 
pl_ts = ggplot() + 
  geom_line(
    data = params$standard_count_data %>% dplyr::select(date, intake),
    aes(x = date, y = intake)
  ) + theme_minimal() +
      scale_x_datetime("Date", 
        limits = c(lubridate::ymd_hms("2020-05-01 00:00:00"), 
                   lubridate::ymd_hms("2021-04-01 00:00:00"))) +
      scale_y_continuous("Daily detainee intake (count)")
pl_hist = ggplot() +
  geom_histogram(
    data = params$standard_count_data %>% dplyr::select(date, intake),
    aes(x = intake), bins = 100
  ) + theme_minimal() +
      scale_x_continuous("") +
      scale_y_continuous("count") +
      coord_flip()
pl = gridExtra::arrangeGrob(pl_ts, pl_hist, 
  layout_matrix = matrix(data = c(1, 1, 1, 1, 2), nrow = 1, ncol = 5)) 
grid::grid.draw(pl)
```


A quarter of all individuals are released within a week of initial detention
and half of all individuals are released within 17 days.  An addtional quarter
remain detained up to three months and the longest detentions recorded in our 
dataset were over a year in duration.  This limits the opportunity for any
particular individual to be affected by within-jail transmission but does
increase the likelyhood that cases originating in the jail will be spread
in the community once detainees are released.  It also increases the number of 
potentially exposed individuals as the jail population is relatively constant
despite the fast turnover.


```{r duration-of-detention, include = TRUE, fig.cap = "Empirical cumulative distribution for the duration of detention, x-axis is the duration of detention (from the MRSA data) and y-axis is the percentile for that duration.  It takes about 87 days to cover the 75% of detentions and there's a long tail up to two years."}
pl = ggplot(
  data = params$duration_of_detention_quantiles,
  aes(x = duration_of_detention, y = (percentile * 100))
) + geom_line() +
    geom_text(
      data = params$duration_of_detention_quantiles %>%
        dplyr::filter(percentile %in% c(0.25, 0.5, 0.75)),
      aes(x = duration_of_detention + 10, y = (percentile * 100) - 5, 
          label = round(duration_of_detention))) +
    geom_point(
      data = params$duration_of_detention_quantiles %>%
        dplyr::filter(percentile %in% c(0.25, 0.5, 0.75, 0.9)),
      aes(x = duration_of_detention, y = (percentile * 100)),
      color = 'red') +
    scale_x_continuous("Duration of detention (days)") +
    scale_y_continuous("Percentile (%)") +
    theme_minimal()
print(pl)
```

```{r occupancy-data, include = FALSE}
total_occupancy = params$standard_occupancy_data %>%
  dplyr::group_by(date) %>%
  dplyr::summarize(count = sum(count)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    count = dplyr::if_else(is.na(count), 
      (dplyr::lag(count) + dplyr::lead(count))/2, count),
    lead_count = dplyr::lead(count)
  )
intake_discharge = total_occupancy %>% 
  dplyr::mutate(date = lubridate::as_date(date)) %>%
  dplyr::left_join(
    y = params$standard_count_data %>%
      dplyr::select(date, intake, isolated_count, cumulative_covid_count, covid_count) %>%
      dplyr::mutate(date = lubridate::as_date(date)), 
    by = 'date') %>%
  dplyr::mutate(
    intake_imputed = is.na(intake),
    intake = dplyr::case_when(
      !intake_imputed ~ intake,
      intake_imputed ~ sample(
        x = intake %>% purrr::discard(is.na),
        size = dplyr::n(),
        replace = TRUE)
      ),
    discharge_count = -(lead_count - (count + intake)),
    covid_discharge_count = discharge_count * (isolated_count / count) 
  )
early_occupancy = total_occupancy %>% 
  dplyr::arrange(date) %>% 
  dplyr::pull(count) %>%
  `[`(1)
lowest_occupancy = total_occupancy %>%
  dplyr::arrange(count) %>% 
  `[`(1,) %>%
  dplyr::mutate(date = lubridate::as_date(date))
dates = c(4500, 5000, 5500) %>%
  rlang::set_names() %>% 
  purrr::imap( ~ total_occupancy %>% 
    dplyr::filter(date > lowest_occupancy$date + lubridate::days(5)) %>%
    dplyr::filter(date < lubridate::ymd_hms("2020-12-01 00:00:00")) %>%
    dplyr::mutate(distance = abs(count - .x)) %>%
    dplyr::arrange(distance) %>%
    dplyr::pull(date) %>%
    dplyr::first() %>%
    lubridate::as_date()
  )

```

With these patterns of intake and discharge, the population of the jail fell
from an early high of `r paste0(substr(early_occupancy, 1, 1), ",",
substr(early_occupancy, 2, nchar(early_occupancy)))` in 
`r lubridate::as_date(min(total_occupancy$date, na.rm = TRUE))` to 
a low of `r lowest_occupancy$count` in `r lowest_occupancy$date` and quickly
rose again to 4,500 around `r dates[['4500']]`, 5,000 around `r dates[['5000']]` and the 
original high occupancy of circa 5,500 by `r dates[['5500']]` only
`r difftime(dates[['5500']], lowest_occupancy$date, units = 'days')` days after the
low point.  This pattern can also be seen in the percentage occupancy plots 
which demonstrate that the jail was only able to keep some units at 50%
capacity.  

```{r occupancy-time-series, layout="l-body-outset", include = TRUE, fig.cap="While the depopulation of the jail early in the pandemic resulted in decreased density within the jail, later increases returned the population to near original occupancy."}
m = 4
pl_occ_ts = ggplot() +
  geom_ribbon(
    data = intake_discharge %>% dplyr::mutate(date = lubridate::as_datetime(date)),
    aes(x = date, ymin = 0, ymax = intake),
    fill = 'blue', alpha = 0.5
  ) + 
  geom_ribbon(
    data = intake_discharge %>% dplyr::mutate(date = lubridate::as_datetime(date)),
    aes(x = date, ymin = count - discharge_count, ymax = count),
    fill = 'orange', alpha = 0.5
  ) + 
  geom_line(
    data = total_occupancy,
    aes(x = date, y = count)
  ) +
  geom_point(
    data = dplyr::bind_rows(
      lowest_occupancy,
      tibble::tibble_row(date = dates[['4500']], count = 4500),
      tibble::tibble_row(date = dates[['5000']], count = 5000),
      tibble::tibble_row(date = dates[['5500']], count = 5500)
    ) %>% dplyr::mutate(date = lubridate::as_datetime(date)),
    aes(x = date, y = count),
    colour = 'red') +
  geom_text(
    data = dplyr::bind_rows(
      tibble::tibble_row(date = lowest_occupancy$date + lubridate::days(9*m),
                         count = lowest_occupancy$count - 8*m),
      tibble::tibble_row(date = dates[['4500']] + lubridate::days(9*m), count = 4500 - 8*m),
      tibble::tibble_row(date = dates[['5000']] + lubridate::days(9*m), count = 5000 - 8*m),
      tibble::tibble_row(date = dates[['5500']] + lubridate::days(9*m), count = 5500 + 35*m)
    ) %>% dplyr::mutate(date = lubridate::as_datetime(date)),
    aes(x = date, y = count - 20, label = lubridate::as_date(date))
  ) + theme_minimal() + 
      scale_x_datetime("Date") +
      scale_y_continuous("Jail Detainee Population (count)", limits = c(0, 5800))
print(pl_occ_ts)
```

At the same time COVID-19 cases continue to enter and exit the jail and
propagate within the facility.  

```{r testing-and-control, layout="l-body-outset", include = TRUE, fig.cap="Grey ribbon represents the cases isolated within the facility, orange ribbon represents the cases assumed released into the community as a proportion of regular releases, and the red line is the cumulative number of cases released into the community up to a time."}
id = intake_discharge %>% dplyr::filter(!intake_imputed)
ggplot() + 
  geom_ribbon(
    data = id,
    aes(x = date, ymin = 0, ymax = isolated_count),
    alpha = 0.5
  ) + 
  geom_line(
    data = id,
    aes(
      x = date, 
      y = cumsum(covid_discharge_count)
    ),
    color = 'red', alpha = 0.5
  ) +
  geom_ribbon(
    data = id,
    aes(x = date, ymin = 0, ymax = covid_discharge_count),
    fill = 'orange'
  ) + 
  theme_minimal()


```

# Occupancy

During the period April 2020 to April 2021 occupancy was reduced to 50%
capacity in most units, with the exception of building 8 and the building
8 RTU.  Starting in late 2020 all buildings with the exception of building 
2 had a significant proportion of units above 55% capacity. 

```{r occupancy, include = FALSE}
unit_occupancy = params$standard_occupancy_data %>% 
  dplyr::left_join(
    y = params$standard_capacity_data %>% dplyr::select(building, unit, capacity),
    by = c('unit')) %>%
  dplyr::filter(!is.na(building)) %>% 
  dplyr::mutate(
    percentage = round(count / as.numeric(capacity) * 100),
    building = paste0("Building ", building)
  )

pl = unit_occupancy %>% 
  dplyr::pull(building) %>% 
  unique %>% 
  purrr::discard( ~ .x == 'Building 16') %>%
  purrr::map( ~ ggplot() +
      geom_line(
        data = unit_occupancy %>% dplyr::filter(building == .x), 
        aes(x = date, y = percentage, group = unit),
        alpha = 0.15) + 
      geom_line(
        data = unit_occupancy %>% 
          dplyr::filter(building == .x) %>%
          dplyr::group_by(date, building) %>% 
          dplyr::summarize(
            occupancy = (1 - sum(percentage < 55, na.rm = TRUE)/
                              sum(!is.na(percentage))) * 100
          ) %>% dplyr::ungroup(),
        aes(x = date, y = occupancy),
        color = 'blue') +
      facet_wrap( ~ building) + 
      scale_x_datetime("Date") +
      scale_y_continuous("Occupancy (%)") +
      theme_minimal() +
      coord_cartesian(ylim = c(0, 100))
  )
```

```{r pl-occupancy, include = TRUE, fig.cap="Each faded black line represents the recorded occupancy (as a percentage of capacity) of each unit within a building.  The blue line represents the proportion of units above 55% occupancy." }
o = purrr::map(pl, print)
```

# Intake testing levels

Tests are done on intake but not mandatory, though compliance is consistently
high.

```{r testing-levels, layout="l-body-outset", include = TRUE, fig.cap = "Testing rate (top, black line) and observed positivity rate (bottom, black dots) and estimated (7-day averaged) positivity rate credible intervals (central 50% and central 90%)"}
assumed_positives = 1
assumed_tests = 100

positivity = dplyr::bind_rows(
  tibble::tibble(
    date = lubridate::ymd_hms("2020-05-22 07:00:00"), 
    positives = assumed_positives, tests = assumed_tests,
    cumulative_window_positives = 1, 
    cumulative_window_negatives = 99), 
  params$standard_count_data %>% 
    dplyr::transmute(
      date = date,
      positives = covid_count, 
      tests = test_count) %>%
    dplyr::filter(!is.na(positives), !is.na(tests))
  ) %>%
  dplyr::mutate(
    cumulative_window_positives = positives %>% windowed(8, sum),
    cumulative_window_negatives  = (tests - positives) %>% windowed(8, sum),
    beta_10 = purrr::map2(cumulative_window_positives, cumulative_window_negatives,
        ~ qbeta(p = 0.10, shape1 = .x, shape2 = .y)) %>% 
      purrr::flatten_dbl(),
    beta_25 = purrr::map2(cumulative_window_positives, cumulative_window_negatives,
        ~ qbeta(p = 0.25, shape1 = .x, shape2 = .y)) %>%
      purrr::flatten_dbl(),
    beta_75 = purrr::map2(cumulative_window_positives, cumulative_window_negatives,
        ~ qbeta(p = 0.75, shape1 = .x, shape2 = .y)) %>%
      purrr::flatten_dbl(),
    beta_90 = purrr::map2(cumulative_window_positives, cumulative_window_negatives,
        ~ qbeta(p = 0.90, shape1 = .x, shape2 = .y)) %>%
      purrr::flatten_dbl(),
  )

pl_testing = ggplot() +
  geom_line(
    data = params$standard_count_data,
    aes(x = date, y = proportion_tested),
    color = 'black', alpha = 0.5) +
    theme_minimal() + 
    theme(axis.text.x = element_blank()) +
    scale_x_datetime("") +
    scale_y_continuous("intake testing level (%)")
pl_positivity = ggplot() +
  geom_point(
    data = params$standard_count_data,
    aes(x = date, y = 100 * test_positivity)) +
  geom_ribbon(
    data = positivity,
    aes(x = date, ymin = 100 * beta_10, ymax = 100 * beta_90),
    fill = 'blue', alpha = 0.2) +
  geom_ribbon(
    data = positivity,
    aes(x = date, ymin = 100 * beta_25, ymax = 100 * beta_75),
    fill = 'blue', alpha = 0.2) +
  theme_minimal() +
  scale_x_datetime("date") +
  scale_y_continuous("test positivity (%)")
pl = gridExtra::arrangeGrob(pl_testing, pl_positivity, 
  layout_matrix = matrix(data = c(1, 2, 2), nrow = 3, ncol = 1)) 
plot.new()
grid::grid.draw(pl)
```

# Movement


```{r observed-movement-counts, layout="l-body-outset", include = TRUE}
params$building_count %>%
  dplyr::rename(`row sums` = row_sums) %>%
  knitr::kable(caption = "Observed movement counts between buildings.")
```

```{r observed-movement-percentage, layout="l-body-outset", include = TRUE}
params$building_percentage %>%
  dplyr::rename(`row sums` = row_sums) %>%
  knitr::kable(caption = "Observed percentages of movement between buildings.")
```

```{r observed-movement-within-building, layout='l-body-outset', include = TRUE}
#unit_pl_data = params$building_count$building %>% purrr::map(
#  ~ params$unit_mmo %>%
#    dplyr::filter(building == .x, last_stable_building == .x)) %>%
#  purrr::discard(~ nrow(.x) == 0) %>% 
#  purrr::discard(~ all(.x$row_sums == 0)) %>%
#  purrr::discard(~ all(.x$row_sums == 0)) %>%
#  purrr::discard(~ all(.x$building %in% c("Intake", "01", "03", "05", "16", "RCDC")))
#
#unit_grid_pl = unit_pl_data %>% 
#  purrr::map(~ ggplot(data = .x) + geom_raster(aes(x = unit, y = last_stable_unit, fill = percentage)))
```

```{r unit-transition-pl, layout="l-body-outset", include = TRUE, fig.cap = "Units ordered on each axis grouped within building, light blue blocks are buildings, zeros are ignored and brighter blue corresponds to more moves between units."}
print(params$unit_transition_plot)
```
